\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-07-01 exps.tex
Included all changes by Prof. Salcic

\section{Experimental results}
\label{sec:experimental-results}

@@

\begin{figure}
\centering
	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
\caption{Evaluation of logical delays \emph{d} for on various benchmark programs}\label{fig:comparison}
\end{figure}
%
In this section, we present a set of SystemJ benchmark programs in which
real-time requirements should be met.
We have carried out a set of experiments to obtain BCRT and WCRT of the
benchmark programs on the execution platform called Java Optimized
Processor(JOP) \cite{jop:jnl:jsa2007}. JOP is a hardware implementation of the
JVM written in VHDL which enables real-time execution of Java programs by
translating Java bytecodes into a sequence of JOP's native instructions called
\emph{microcode} which is time-predictable. As SystemJ's default compilation
target is Java source code, JOP is an excellent platform which enables us to
analyze timing properties of SystemJ program.

@@

By utilizing in-house tools in conjunction with Worst-Case-Execution-Time (WCET)
analyzer \cite{jop:jnl:jsa2007} provided by JOP tools, we were able to estimate
BCRT and WCRT of the following benchmark programs: (full-name?) HRTCS \cite{},
Stepper motor controller (Motor) \cite{}, Robot and Access Environment Control
System (AECS) \cite{}. We have also added a set of customized hardware counters
to the original JOP design in order to collect numbers of clock cycles required
for logical ticks of given SystemJ program. All examples except Robot are
existing applications translated into pure SystemJ programs. The Robot example
models a combination of robotic arms and conveyors which splits goods based on
their volumetric sizes. Finally, garbage collector (GC) was disabled in all
experiments to avoid unwanted side-effects such as lengthening clock-domain
logical tick times. Whenever the GC triggers, experiment terminates. BCRT and
WCRT were then calculated based on data collected until that point.

As one can see in Figure \ref{fig:comparison}, BCRT and WCRT were calculated for
all examples. Given the real-time ranges \mbox{\texttt{delay [M..N]}} provided
by programmers, required logical delays \emph{d} were determined by applying the
Algorithm \ref{alg:1}. In HRTCS, real-time delay of the range 50.3..200.3 (ms)
is only satisfied when \emph{d} is between 1506 and 1787 (not shown in Figure).
It is then compiler's job to statically choose one of the values between this
range.

However, it is not the case that given real-time delays in a particular
application are always realizable on JOP. For instance, all other examples other
than HRTCS are required to relax the upper bound for a given real-time delay in
order to obtain a feasible number of \emph{d}. Therefore, Algorithm \ref{alg:2}
is applied to those examples to determine the minimum upper real-time bound
\emph{N} such that there exist a solution for \emph{D} in Algorithm \ref{alg:1}.
Unlike in HRTCS, however, this algorithm gives a unique number for logical tick
delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).
The compiler, therefore, will choose \emph{d} based on the newly computed \emph{N} for
the examples Robot, Motor and AECS.

@@

% As one can see in Figure \ref{fig:comparison}, BCRT and WCRT of overall programs
% are smaller when they run on TP-JOP. For example, WCRT and BCRT of HRTCS are
% \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028,
% 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other hand, required
% logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 -
% \(\times\)5.9 and \(\times\)30.9 in Motor and Robot respectively. It is expected
% as their BCRT and WCRT differences are bigger on TP-JOP.
% This also led to greater increase in minimum upper real-time bounds \emph{N} for
% TP-JOP in every example.
% In AECS, there are two clock-domains using delay constructs that each has their
% own BCRT and WCRT. Again, individual BCRT and WCRT is smaller whereas \emph{d}
% is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT
% of any clock-domains are invariant of \emph{d} as explained in Section
% \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in
% our benchmark programs.

>>>>

\section{Experimental results}
\label{sec:experimental-results}
% <<<<<<< HEAD
% \begin{table}[t]
%   \centering
%   \twocolumn[

%   \begin{@twocolumnfalse}
%     \caption{Benchmark programs using \texttt{delay} constructs} \label{table:benchmark}
%     \scalebox{0.85}{
%       \begin{tabular}{|c| c | c | c | c | c | c | c |}
% 	\cline{4-8}
% 	\multicolumn{3}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
% 	\multirow{12}{*}{JOP} 		& \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{2-8}   
%         & \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{2-8} 
%         & \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{3-8} 
%         &			   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{3-8} 
%         &			   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
%         &			   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
%         &			   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
%         &			   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}   
%         & \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8} 
%         & \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8} 
%         &			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8} 
%         &			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8} 
%         &			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8} 
%         &			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
%         & \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8} 
%         &			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
% 	\hline
%       \end{tabular}
%     }
%   \end{@twocolumnfalse}
%   ]

% \end{table}
% In this section, we present a set of SystemJ benchmark programs in which
% real-time requirements should be met.  We have carried out a set of
% experiments to obtain BCRT and WCRT of the benchmark programs on two
% execution platforms called Java Optimized Processor(JOP)
% \cite{jop:jnl:jsa2007} and TP-JOP \cite{6119095}. JOP is a hardware
% implementation of the JVM which enables real-time execution of Java
% programs by translating Java bytecodes into a sequence of JOP's native
% instructions called \emph{microcode} which is time-predictable. As
% SystemJ's default compilation target is Java source code, JOP is an
% excellent platform which enables us to analyze timing properties of
% SystemJ program. On the other hand, there is also an option where
% compiled code could be more tightly coupled to specific platforms such
% as TP-JOP, which execute SystemJ's kernel statements more
% efficiently. By utilizing our internal tools in conjunction with
% Worst-Case-Execution-Time (WCET) analyzer \cite{jop:jnl:jsa2007}
% provided by JOP tools, we were able to estimate BCRT and WCRT of our
% benchmark programs: (full-name?) HRTCS, Stepper motor controller (Motor)
% \cite{}, Robot and Access Environment Control System (AECS) \cite{}.

% % HRTCS, as already illustrated in Section \ref{sec:intr-motiv}, tests
% % for an ability of one's responsiveness by generating green lights
% % between the real-time range \emph{M - N}. For AECS \ref{} we have
% % replaced the external timer with our delay constructs. Stepper motor
% % controller is originally introduced in \cite{Bourke2009a} which
% % converted into SystemJ.

% As one can see in Table \ref{table:benchmark}, BCRT and WCRT of overall
% programs are smaller when they run on TP-JOP. For example, WCRT and BCRT
% of HRTCS are \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for
% TP-JOP (0.0028, 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the
% other hand, required logical delays \emph{d} are generally bigger on
% TP-JOP e.g. \(\times\)4.5 - \(\times\)5.9 and \(\times\)30.9 in Motor
% and Robot respectively. It is expected as their BCRT and WCRT
% differences are bigger on TP-JOP.  This also led to greater increase in
% minimum upper real-time bounds \emph{N} for TP-JOP in every example.  In
% AECS, there are two clock-domains using delay constructs that each has
% their own BCRT and WCRT. Again, individual BCRT and WCRT is smaller
% whereas \emph{d} is bigger on TP-JOP. One important property to note
% here, is that BCRT and WCRT of any clock-domains are invariant of
% \emph{d} as explained in Section \ref{sec:intr-real-time}. It is our
% fundamental assumption to find \emph{d} in our benchmark programs.
% =======

@@

\begin{figure}
\centering
	\includegraphics[trim=2cm 2cm 2cm 2cm,scale=0.5]{table2}
        \caption{Evaluation of logical delays \emph{d} for on various
          benchmark programs}\label{fig:comparison}
\end{figure}
%
We have carried out a number of experiments on a set of applications
with real-time constraints. The benchmark set is shown in
Figure~\ref{fig:comparison}. HRCTS is the human response system
described in Figure~\ref{fig:1}. AECS (Access and Environmental Control
System) is an application that controls an intelligent room environment
described in~\cite{aecs_ispa}. Motor is a stepper motor controller
from~\cite{Bourke2009a}. Robot is an industrial automation application
that consists of two manufacturing belts and a robotic arm that splits
goods according to their volumes. HRCTS consists of 2 CDs, but only one
with real-time delays. AECS also has 2 CDs; CD1 consist of a single
delay statement, while CD2 has 2 delay statements. Robot and Motor both
are synchronous programs with 1 and 5 delay statements,
respectively. All these examples are written in the SystemJ
language. Finally, except for HRCTS all other examples have exact (and
deterministic) delay requirements. These delay requirements are shown in
Figure~\ref{fig:comparison} (\texttt{M..N}).

% In this section, we present a set of SystemJ benchmark programs in which
% real-time requirements should be met. We have carried out a set of
% experiments to obtain BCRT and WCRT of the benchmark programs on the
% execution platform called Java Optimized Processor(JOP)
% \cite{jop:jnl:jsa2007}. JOP is a hardware implementation of the JVM
% written in VHDL which enables real-time execution of Java programs by
% translating Java bytecodes into a sequence of JOP's native instructions
% called \emph{microcode} which is time-predictable. As SystemJ's default
% compilation target is Java source code, JOP is an excellent platform
% which enables us to analyze timing properties of SystemJ program.

@@

We use the \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}, to carry out all our experiments. JOP
implements the JVM specification in hardware directly, and has been
shown to be real-time amenable. We statically estimate the WCRT and BCRT
of these benchmark applications using the real-time analysis tools
provided by the JOP design framework. We disable the \textit{Garbage
  Collector} (GC) in all experiments to avoid unwanted side-effects such
as lengthening clock-domain logical tick times.

As one can see in Figure \ref{fig:comparison}, BCRT and WCRT were
calculated for all examples. Given the real-time \mbox{\texttt{delay
    (M..N)}} provided by programmers, required logical delay \texttt{d}
was determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time
delay of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is
between 1506 and 1787 (as shown previously in
Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
then compiler's job to statically choose one of the values between this
range.

The upper bound \texttt{N} needed to be relaxed for all examples other
than HRCTS to obtain a feasible \texttt{d}. Therefore, Algorithm
\ref{alg:2} is applied to those examples to determine the minimum upper
real-time bound \emph{N} such that there exist a solution for \emph{D}
in Algorithm \ref{alg:1}.  Unlike in HRTCS, however, this algorithm
gives a unique number for logical tick delay \emph{d} such that \emph{d}
= Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler, therefore, will
choose \emph{d} based on the newly computed \emph{N} for the examples:
Robot, Motor and AECS.

@@

% As one can see in Figure \ref{fig:comparison}, BCRT and WCRT of overall programs
% are smaller when they run on TP-JOP. For example, WCRT and BCRT of HRTCS are
% \(\times\)11.9 and \(\times\)3.4 smaller, respectively, for TP-JOP (0.0028,
% 0.0329 ms) compared to JOP (0.0334, 0.1120 ms). On the other hand, required
% logical delays \emph{d} are generally bigger on TP-JOP e.g. \(\times\)4.5 -
% \(\times\)5.9 and \(\times\)30.9 in Motor and Robot respectively. It is expected
% as their BCRT and WCRT differences are bigger on TP-JOP.
% This also led to greater increase in minimum upper real-time bounds \emph{N} for
% TP-JOP in every example.
% In AECS, there are two clock-domains using delay constructs that each has their
% own BCRT and WCRT. Again, individual BCRT and WCRT is smaller whereas \emph{d}
% is bigger on TP-JOP. One important property to note here, is that BCRT and WCRT
% of any clock-domains are invariant of \emph{d} as explained in Section
% \ref{sec:intr-real-time}. It is our fundamental assumption to find \emph{d} in
% our benchmark programs.
% >>>>>>> 4bffc4b1a43388f5ebf6cc737b9d3b246ee77dc7

### 2013-07-01 introduction.tex
Included all changes by Prof. Salcic

Reactive languages~\cite{gber05,amal10} are a class of programming
languages used for designing and implementing \textit{reactive} systems,
which continuously respond to input from their environment. These
languages have been successfully used in programming a plethora of
systems such as fly by wire in Airbus~\cite{eairbus}, security
surveillance systems~\cite{amal121}, etc. Usually the working
environment of these reactive systems is also real-time. Yet, these
languages do not support describing real-time as first class language
constructs. For example, one cannot describe a simple real-time delay
\texttt{delay 0.2 ms} in these languages. These languages are based on
formal semantics, essential for the verification of functional
correctness properties of the developed programs, but leave
non-functional properties such as timing behavior as an implementation
detail~\cite{ringler00,boldt07}. Arguably, rightly so, because physical
time cannot be incorporated without information about the underlying
execution platform. But, one can still reason in terms of
\textit{logical} time. These languages support discrete \textit{logical}
clock rather than a continuous clock. The period of the discrete logical
clock (also referred to as a \textit{logical tick}) is determined by the
responsiveness of the program to the input signals and as opposed to a
discrete \textit{physical} clock, is elastic. Although this timing model
has worked well for these languages in designing discrete systems, there
is obviously a need to introduce continuous time since often
implementation models are in the continuous time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.

@@

The fundamental contribution of this paper is the inclusion of real-time
constructs in GALS (and its subset synchronous) reactive
languages. Moreover, our contributions can be further refined as
follows:

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
\item We consider the property of \textit{non maximal-parallelism},
  i.e., we consider scheduling on constrained resources to be an
  integral part of the solution. To our knowledge we are the very first
  to do so.
\item Our solution does not change the functional or timing semantics of
  GALS and its subset synchronous programs, in fact, our solution does
  not require one to change the mid-end, the back-end or the related
  optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages.
\end{enumerate*}

The rest of the paper is arranged as follows...

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc.  Usually these systems also need to meet
real-time constraints enforced by the environment.  Yet, the languages
used to program these systems do not support describing real-time
constructs as first class language constructs.  For example, one cannot
describe a simple real-time delay (postponement) of an operation for 0.2
ms. These languages are based on formal semantics, essential for the
formal reasoning about and verification of correctness of functional
properties of the developed programs, but leave nonfunctional properties
such as timing behavior as an implementation
detail~\cite{boldt07}. Arguably, rightly so, because physical time
cannot be incorporated without information about the underlying
execution platform.  But, one can still reason in terms of logical
time. These languages support discrete logic clock rather than a
discrete physical (real-time) clock. The real-time period of the
discrete logic clock (also referred to as a logic tick) is not fixed and
it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logic clock is elastic. Although the timing
model with logic clock has worked well for reactive synchronous
languages in designing discrete systems, there is a need to introduce
real-time since, often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as the ﬁrst class language
constructs create additional opportunity for formally verifying
functional and non-functional requirements before system deployment.

@@

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delays model relays on
the use of relative instead of absolute time, i.e., a delay in selected
time units is counted from the currently executing SystemJ statement and
(3) delays allow certain behaviours to having real-time features, while
the other behaviours use only logical time and (4) real-time delay does
not affect the model of logical time in SystemJ program as long as the
amount of delay is within boundaries that can be determined statically
by program analysis. All these as well as semantics of the delay
construct are discussed and illustrated in the remaining part of the
paper. Our contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reduced power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify a
  single \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range, of real-time values for a delay
  statement, e.g., Figure~\ref{fig:1}.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time delays in
the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:experimental-results} gives the experimental
results for a set of real-time
applications. Section~\ref{sec:related-work} positions this work in
relation to other approaches to introducing real-time. Finally, we
conclude in Section~\ref{sec:concl-future-work}.

### 2013-07-01 motivation.tex
Included all changes by Prof. Salcic

\section{Related work}
\label{sec:related-work}

>>>>

\subsection{Motivating Example}
\label{sec:motivating-example}

% system{
%  interface{
%   input signal TOUCH;
%   output signal GREEN_LIGHT;
%   output char channel S;
%   input char channel R;
%  }
\begin{figure}[t!]
% \begin{center}
\begin{minipage}{5cm}
  \begin{scriptsize}

\begin{verbatim}
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;}
    || // synchronous parallel operator
    {
    //exit after any 
    // where from 50.3 to 200.3 ms.
    delay (50.3 .. 200.3 ms);
    exit(TRAP); 
    }
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S;}
  // do something with S
}
\end{verbatim}
  \end{scriptsize}
\end{minipage}
\caption{An example human responsiveness system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a simple machine used to test human response time and collect
this data for future research purposes. The machine switches on a green
light on a touch screen for anywhere between \texttt{50.3ms} to
\texttt{200.3ms}, if the user can touch this green light then she is
successful and an integer \texttt{1} is sent to a database, if
unsuccessful a \texttt{0} is sent instead, Different statistics and
analyses on human responsiveness can be based on collected data.

The SystemJ psuedocode~\cite{amal10} code for such a system is shown in
Figure~\ref{fig:1}. We use the SystemJ language in this paper first of
all, because being a \textit{Globally Asynchronous Locally Synchronous}
(GALS) language it helps us easily describe a larger class of systems
(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
allows us to explore the marriage of real-time and logical time, which
is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of
\texttt{50.3ms} to \texttt{200.3ms} a positive response is sent to the
second clock-domain. All the SystemJ programming constructs used to
implement this system are described in~\cite{amal10} and
Table~\ref{tab:1} in Section~\ref{sec:background}. The delay
speciﬁcation, which models the non-deterministic delay between 50.3ms
and \texttt{200.3ms} is not supported in the current language and cannot
be compiled with the current SystemJ compiler.

### 2013-07-01 related_work.tex
Included all changes by Prof. Salcic

A small number of studies have been performed when it comes to
integrating real-time directly as a programming construct in GALS (and
its subset synchronous) reactive languages. The most prominent works in
this area are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. do a much better job of introducing real-time
as first class constructs in the Esterel language. They like us provide
real-time \texttt{delay} as first class Esterel programming constructs,
and translate them into Esterel kernel constructs. But, unlike us they
do not translate delays into \texttt{pause} constructs
directly. Instead, logical ticks are generated by using abstract notion
of \texttt{event} and \texttt{sample} platform dependent timers. This
notion makes it a complex and inflexible solution since the number of
platform timers with certain resolutions need to be determined and
present on the system for the solution to be realizable. Moreover,
non-deterministic real-time delays and integration with non-maximal
parallelism is not studied at all by them. Other works such as
Quartz~\cite{glog02} also incorporate delay statements in Esterel using
\texttt{pause} constructs, but again, non-maximal parallelism,
non-deterministic delays are not studied. Moreover, Quartz is targeted
at studying timing properties using model checkers rather than
implementation, like us.

>>>>

A small number of studies have been performed when it comes to
integrating real-time directly as a programming construct in GALS (and
its subset synchronous) reactive languages. The most prominent works in
this area are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
delay in logical ticks \textit{after} allocation and scheduling, it
gives the compiler developers the chance to optimize for many different
criteria such as computation time, energy, power, etc, without worrying
about or violating the real-time delay specification. Other works such
as Quartz~\cite{glog02} also incorporate delay statements in Esterel
using \texttt{pause} constructs, but again, non-maximal parallelism and
non-deterministic delays are not studied. Moreover, Quartz is targeted
at studying timing properties using model checkers rather than
implementation, like us.
